Enabling Ambient Backscatter \\Using Low-Cost Software Defined Radio

### 2017-04-07 paper.tex
Added introduction
Some cleanup in the evaluation section

%% bare_conf.tex
%% V1.4b
%% 2015/08/26
%% by Michael Shell
%% See:
%% http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
%% conference paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/pkg/ieeetran
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall the IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. The IEEE's font choices and paper sizes can   ***
% *** trigger bugs that do not appear when using other class files.       ***                          ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

\documentclass[conference]{IEEEtran}
% Some Computer Society conferences also require the compsoc mode option,
% but others use the standard conference format.
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[conference]{../sty/IEEEtran}

@@

\newcommand{\cp}[1]{\footnote{{\bf Carlos: #1}}}

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Enabling Ambient Backscatter Using Low-cost Software-defined-radio}

@@

\hfill April 7, 2017

@@

\section{Evaluation}

\subsection{Signal Strength Variation within the City}
To observe the signal strength variation within the city the \textit{RTL-SDR} has been used as well.

@@

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
\bibliography{paper}
\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
%\begin{thebibliography}{1}
%
%\bibitem{github:librtlsdr}
%Steve Markgraf et. al. \emph{librtlsdr}. \url{https://github.com/steve-m/librtlsdr}. Github: 2012-2017.
%
%\bibitem{rafael:r820t}
%\emph{High Performance Low Power Advanced Digital TV Silicon Tuner}. R820T. Rev 1.2. Rafael Micro. November 2011.
%
%\bibitem{book:pozar}
%D. M. Pozar. \emph{Microwave Engineering}. Fourth Edition. John Wiley \& Sons Inc. 111 River Street, Hoboken. 2012
%
%\end{thebibliography}

% that's all folks
\end{document}

>>>>

\documentclass[conference]{IEEEtran}
\usepackage{booktabs} % For formal tables
\newenvironment{itemize*}{\begin{itemize}}{\end{itemize}}
\usepackage[utf8x]{inputenc}
\usepackage{subfigure}
\usepackage{balance}
\usepackage{comment}
\usepackage{amsmath}
\usepackage{xspace}
\usepackage{caption}
\usepackage{siunitx}
\usepackage{booktabs}
% Some Computer Society conferences also require the compsoc mode option,
% but others use the standard conference format.
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[conference]{../sty/IEEEtran}

@@

\newcommand{\cp}[1]{\footnote{{\bf Carlos: #1}}}
\newcommand{\fakeparagraph}[1]{\vspace{.5mm}\noindent\textbf{#1.}}
\newcommand{\fakepar}[1]{\fakeparagraph{#1}}

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Enabling Ambient Backscatter \\Using Low-Cost Software Defined Radio}

@@

\input{introduction}

@@

\section{Evaluation}
In this section we present our evaluation results. The \textit{RTL-SDR}, combined with intelligent signal processing in \textit{Octave}, are the main tools we have employed.

\subsection{Signal Strength Variation in Space}
\begin{figure}[h]
	\centering
	\begin{minipage}{0.49\columnwidth}
	\includegraphics[width=\columnwidth]{./fig/626mhz_raw}
	\end{minipage}
	\hfill
	\begin{minipage}{0.49\columnwidth}
	\includegraphics[width=\columnwidth]{./fig/626mhz_filtered}
	\end{minipage}
	\caption{Spectrum of the TV signal. Right: Raw spectrum of the TV signal with \ensuremath{f_{\text{center}}=\text{626 Mhz}}. Left: Smoothed spectrum, normalized to maximum average measured in decibel with average shown as red line.}
	\label{fig:tv_record} 
\end{figure}

We have used the textit{RTL-SDR} to observe the space variations of the signal strength. \textit{Octave} scripts have been written to aid in this purpose. These scripts are available online under \url{https://github.com/s3xm3x/RTLSDRspecAn}. The scripts perform a frequency sweep through the desired band. The obtained samples are then transferred from the time domain to the frequency domain using an FFT.

Because the RTL-SDR has a maximum sampling rate of 1.8 MSPs, we are
limited to a maximum simultaneous acquisition band of 900 kHz. Because
of this, we sample the desired frequency band at 900 kHz intervals and
stitch these bands together to form the overall frequency sweep. This
approach is valid only because we are not interested in real-time data.
With this approach we are able to scan a 20 MHz band in roughly two
minutes.\cp{Please confirm these things.}

During the scanning process	the gain is set to 40.2 dB. To get the
signal strength we simply take the average over a 10 MHz band around the
center frequency of the desired TV signal. This approach can be justified by the fact, that DVB-T is specified up to 10 MHz bandwidth. And the local TV signal provider is to its own statements using DVB-T 2. The signal \ensuremath{|H|} is calculated as follows.
\begin{equation}
	|H| = \Biggl| \sum_{n=0}^N FFT\bigl( \Re\{ s_{\text{band,n}}(t) \} \bigr) \Biggr|
\end{equation}     
Where \ensuremath{N} is the total number of FFT bins accumulated over
the entire frequency sweep\cp{Please confirm I got this right}. The FFT
is only carried out over the real values, which are received from the
\textit{RTL-SDR}. \ensuremath{s_{\text{band,n}}(t)} is the signal in the
time domain of one band. There are as many samples taken in one band as
needed for an FFT of size 2048. No window is used when stitching
individual scans together. One can see, the effects of this a little bit, when taking a closer look at the plotted data. Rising and falling edges at the beginning of each band are visible. To get \ensuremath{|H_{tv,avg}|} one has to normalize everything and calculate the power. Before doing that a simple smoothing algorithm (FIR) is applied on \ensuremath{|H|}.
\begin{equation}
	|H_{tv,avg}| = 20 \cdot log (|H|) - 20 \cdot log(|H_0|)
\end{equation}   
where \ensuremath{|H_0|} is a normalization value measured at a certain
point in space. 
\begin{figure}[h]
\centering
\includegraphics[width=0.5\columnwidth]{./fig/haversine}
\caption{Signal strength fading against distance relative to the spot where the maximum signal strength has been captured (TV towers). 474 MHz and 626 MHz belong to a tower in Uppsala, Vedyxa and 594 MHz belongs to a tower in Uppsala, Rickomberga. Distances have been calculated with the Haversine formula.}
\label{fig:haversine}
\end{figure}

@@

% references section
\bibliographystyle{IEEEtran}
\bibliography{paper}

\end{document}
